---
title: "Tutorial"
author: "Salvatore Milite"
date: "13/06/2022"
output: html_document
---

## Fitting a linear model


```{r, message=FALSE, warning=FALSE, results='hide'}
library(Seurat)
library(SeuratData)
library(reticulate)
library(dplyr)
library(SVDE)


reticulate::use_condaenv("scvi", required = T)

# Get the dataset online
# InstallData("pbmc3k")

data("pbmc3k")

# Count matrix

input_matrix <- Assays(pbmc3k, slot = "RNA")
input_matrix <- input_matrix@counts[rowSums(input_matrix@counts > 0) > 10,]
input_matrix <- input_matrix %>% as.matrix()

ncounts <- pbmc3k$nCount_RNA / mean(pbmc3k$nCount_RNA)
```
The easiest way to use the regression is to just regress out the library size and normalize the data (the process is similar to what scTransform version 2 does but the interpretation is a bit different, more details in the paper)

```{r}
# Basically equivalent to Pearson residuals in scTransform

model_matrix <- model.matrix(~ 1, pbmc3k@meta.data) # we just need an intercept 
head(model_matrix)


model_intercept <- fit_linear_model(input_matrix,model_matrix,ncounts, method_specific_args = list(batch_size = 1000L))

```

And then you can use the coefficients for PCA and the usual downstream analysis

```{r, warning=FALSE, message=FALSE}
pbmc3k_norm <- pbmc3k

# by default we calculate in python the residuals w.r.t. all the coefficients
pbmc3k_norm@assays$RNA@data <- model_intercept$residuals %>% t()

pbmc3k_norm <- FindVariableFeatures(pbmc3k_norm, selection.method = "vst", nfeatures = 2000, verbose = FALSE)

top10 <- head(VariableFeatures(pbmc3k_norm), 10)
plot1 <- VariableFeaturePlot(pbmc3k_norm)
LabelPoints(plot = plot1, points = top10, repel = TRUE)


pbmc3k_norm <- ScaleData(pbmc3k_norm, verbose = FALSE)

pbmc3k_norm <- RunPCA(pbmc3k_norm, verbose = FALSE)

ElbowPlot(pbmc3k_norm)

pbmc3k_norm <- FindNeighbors(pbmc3k_norm, dims = 1:10, verbose = FALSE)
pbmc3k_norm <- FindClusters(pbmc3k_norm, resolution = 0.5, verbose = FALSE)
pbmc3k_norm <- RunUMAP(pbmc3k_norm, dims = 1:10, verbose = FALSE)

DimPlot(pbmc3k_norm, reduction = "umap",group.by = "seurat_annotations")


```

## Simple test


```{r}
pbmc3k@meta.data$seurat_annotations <- as.character(pbmc3k@meta.data$seurat_annotations) 

#NA as a class
pbmc3k@meta.data$seurat_annotations[is.na(pbmc3k@meta.data$seurat_annotations)] <- "Null"

# No intercept model
model_matrix <- model.matrix(~ seurat_annotations - 1, pbmc3k@meta.data) # we just need an intercept 

default_args <- list(steps = 350L, lr = 0.1,
   gamma_lr = 0.1,
  cuda = FALSE, jit_compile = FALSE,
  full_cov = TRUE, batch_size = 10240L, 
  prior_loc = 0.1, 
  theta_bounds = c(1e-6, 1e6),
  init_loc = 0.1, init_theta = 1000)


default_args$lr <- 0.01
default_args$gamma_lr <- 0.5
default_args$batch_size <- 1000L

model_simple <- fit_linear_model(input_matrix,model_matrix,ncounts, method_specific_args =default_args)

test <- perform_test(model_simple,input_matrix = input_matrix, model_matrix = model_matrix, group1 = "B", group2 = "CD8")

DotPlot(pbmc3k %>% subset(seurat_annotations %in% c("B", "CD8 T")), features = test$gene[1:10], group.by = "seurat_annotations") + Seurat::RotatedAxis()

```


## A more complicated case


## Enforce sparsity


## Run with HMC 

```{r}
model_matrix <- model.matrix(~ 1, pbmc3k@meta.data) # we just need an intercept 

model_intercep_HMC <- fit_linear_model(input_matrix[1:30, 1:200],model_matrix[1:200,,drop = F],ncounts[1:30],  inference_method = "HMC")


```


## Run with MLE 


